Untitled

### 2020-10-19 ms/manuscript.Rmd
Update manuscript
to follow JPR format

---
title: "Untitled"
output: html_document
bibliography: references.bib
editor_options: 
  markdown: 
    wrap: sentence
---

Intergenerational Niche Differentiation in Filmy Ferns (Hymenophyllaceae)^1^

Joel H. Nitta^2,5,6^, James E. Watkins, Jr.^3^, N. Michele Holbrook^2^, Ravahere Taputuarai^4^, Tristan W. Wang^2^, and Charles C. Davis^2^

Running Head: Intergenerational Niche Differentiation in Filmy Ferns

**PREMISE OF THE STUDY**: Desiccation tolerance was a key trait allowing plants to colonize land.
However, little is known about the transition from desiccation tolerant non-vascular plants to desiccation sensitive vascular ones.
Filmy ferns (Hymenophyllaceae) present a useful system to study how water-stress strategies differ between non-vascular and vascular stages within a single organism, as they have vascularized sporophytes and nonvascular gametophytes that are each capable of varying degrees of desiccation tolerance.

**METHODS**: We surveyed sporophytes and gametophytes of 22 species of filmy ferns on Moorea (French Polynesia).
We used chlorophyll fluorescence to measure desiccation tolerance and light responses.
We sequenced plastid *rbcL* and *trnHpsbA* to identify field-collected gametophytes, and conducted phylogenetically informed analyses to identify differences in physiology between life stages and growth habits.

**KEY RESULTS**: Gametophytes had similar or less desiccation tolerance and lower photosynthetic optima than sporophytes.
Despite their lack of greater physiological tolerances, gametophytes of several species occurred over a wider elevational range than conspecific sporophytes.
Epiphytes were more tolerant of desiccation than terrestrial species.

**CONCLUSIONS**: Our results show that filmy fern gametophytes and sporophytes differ in their physiology and niche requirements, and point to the importance of microhabitat in shaping the evolution of water-use strategies in vascular plants.

**KEYWORDS**: chlorophyll fluorescence; desiccation tolerance; DNA barcoding; filmy ferns; French Polynesia; gametophyte; Hymenophyllaceae; Moorea; photosynthesis

^1^Manuscript received \_\_\_\_ ; revision accepted \_\_\_\_.

^2^Department of Organismic and Evolutionary Biology and Harvard University Herbaria, Harvard University, Cambridge, Massachusetts 02138 USA;

^3^Department of Biology, Colgate University, 13 Oak Drive, Hamilton, New York 13346 USA; and

^4^Association Te Rau Ati Ati a Tau a Hiti Noa Tu, Mahina, Tahiti, French Polynesia

^5^Current address: Department of Botany, National Museum of Nature and Science, 4-1-1 Amakubo, Tsukuba, Ibaraki 305-0005 Japan

^6^Author for correspondence (email: joelnitta\@gmail.com)

@@

Ferns represent an important, understudied group of tracheophytes for investigating the transition from desiccation tolerance to desiccation avoidance in land plants.
Ferns are the only major lineage among land plants with sporophytes and gametophytes that are capable of growing independently from each other.
These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (i.e., they regulate their internal water content), whereas fern gametophytes are much smaller (often *\<*1 cm), lack vascular tissue, and are poikilohydric.
Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT [@WatkinsJr.2007c; @Pittermann2013a].
Thus, ferns present a unique opportunity to observe the transition from desiccation tolerance to desiccation avoidance across independently growing life phases within a single organism.

Filmy ferns (Hymenophyllaceae) represent a particularly interesting case for studying the ecological and evolutionary significance of vegetative DT across life stages in ferns.
Filmy ferns are a large family (ca. 430 spp.; PPG I, 2016) of primarily tropical, leptosporangiate ferns named for their remarkably thin leaf laminae, which are usually only a single cell layer in thickness and lack cuticle or stomata (Fig. 1A--D).
Filmy fern sporophytes are an exception to the 'typical' fern sporophyte described above because they have reverted to poikilohydry: their thin laminae allow for passive, rapid uptake and loss of moisture, thereby maintaining nearly constant equilibrium with atmospheric humidity (Raven, 1999).
Many species of epiphytic filmy ferns have extremely reduced root systems (Schneider, 2000), relying completely upon absorption through leaf laminae for water (Shreve, 1911), although terrestrial species possess true roots and use ground water (Dubuisson et al., 2011).
As would be expected from their poikilohydric nature, DT of varying degrees has been reported in sporophytes of many filmy fern species (see references below).

Filmy fern sporophytes encompass a broad diversity of growth forms (terrestrial, saxicolous, hemi-epiphytic, and holoepiphytic) and morphologies, with rhizome and leaf sizes varying over an order of magnitude (Hennequin et al., 2008; Nitta and Epps, 2009; Dubuisson et al., 2013).
Filmy fern gametophytes are similarly diverse, with morphotypes including filamentous, ribbon, and a combination of these two types (Fig. 1E--H).
While there are no quantitative data about DT in filmy fern gametophytes available, DT has been reported in gametophytes from several other fern lineages (Watkins et al., 2007; Watkins and Cardelús, 2012; Pittermann et al., 2013), and may be expected to occur in filmy ferns as well.
The possibility for varying degrees of vegetative DT in both stages of the life cycle of filmy ferns makes them a potentially useful system to compare how DT is utilized between haploid, non-vascular and diploid, vascular life stages.
Furthermore, the diversity of growth forms and habitats in filmy ferns makes them an ideal group for a comparative study of water-stress strategies (i.e., desiccation tolerance vs. desiccation avoidance) and the correlation of these strategies with ecological niche.

The degree of DT in filmy fern sporophytes occupying different niches has been relatively well studied.
In a series of elegant experiments on Jamaican filmy ferns, Shreve (1911) demonstrated that water-use strategies differ between species occupying contrasting habitats (i.e., terrestrial, low elevation epiphyte, and high elevation epiphyte).
Proctor (2003, 2012) investigated water relations and the ability to recover from desiccation in sporophytes of several filmy ferns, and found a general correlation between high DT and adaptation for high light levels in widespread species vs. low DT and adaptation to low light levels in species occupying more sheltered habitats.
In the only study to our knowledge including in situ physiological observations of hymenophyllaceous gametophytes, Johnson et al. (2000) found that gametophytes of *Vandenboschia speciosa* (Willd.) G.
Kunkel (= *Trichomanes speciosum* Willd.) are adapted for extremely low light environments, and operate at lower light levels than conspecific sporophytes.
Parra et al. (2009) and Saldan˜a et al. (2014) investigated the relationship between physiology (photosynthetic rates and water relations) and vertical stratification in several epiphytic filmy fern species (primarily genus *Hymenophyllum*) along tree trunks in Chile.
They found that species occurring at greater heights in the trees tended to have greater DT, higher photosynthetic capacity (A~max~), and lower rates of evapotranspiration.
A recent study using two species of *Hymenophyllum* demonstrated that DT in filmy ferns likely operates via a constitutive mechanism, which is more similar to that of bryophytes than other drought-tolerant vascular plants, thus allowing filmy ferns to adjust to dry conditions extremely rapidly (Cea et al., 2014).

Despite this considerable interest in filmy fern ecophysiology, with the sole exception of Johnson et al. (2000), no studies have compared physiology between gametophyte and sporophyte phases, and we are aware of none that have done so across multiple species.
Furthermore, to our knowledge, no physiological studies of filmy ferns have accounted for their phylogenetic relationships, and we lack an understanding of how physiological strategies have evolved in this clade.
In addition to DT, photosynthetic optima (adaptation to different light levels) are likely to be important in determining the niches of life stages and species of filmy ferns (Proctor, 2003; Parra et al., 2015; Flores-Bavestrello et al., 2016).
Terrestrial and epiphytic species experience different levels of photosynthetic radiation and desiccation, and light intensity varies with height along the host plant in epiphytes (Benzing, 1990; Watkins and Cardelús, 2009).
In a recent study (Testo et al., unpublished), the combined effect of desiccation and light stress was found to have a greater impact on filmy fern physiological performance than either of these stressors alone, suggesting the need to integrate studies of DT with photosynthesis to gain a full picture of physiological adaptation in filmy ferns.
Here, we leverage a recently developed DNA barcoding system (Nitta et al., 2016) to test several hypotheses related to DT and photosynthesis in a comparative phylogenetic framework including both filmy fern sporophytes and gametophytes: 1.) Epiphytic filmy ferns have higher DT and photosynthetic optima than non-epiphytic species.
2.) Amongst epiphytic filmy ferns, those occurring at high elevation have higher DT and photosynthetic optima than those at low elevation.
3.) For filmy ferns with gametophytes that occur over a wider range than sporophytes, the gametophyte has higher DT and photosynthetic optima than the sporophyte.

MATERIALS AND METHODS

***Study system***---We selected the island of Moorea, French Polynesia (17*◦*32' S, 149*◦*50' W) for our survey of filmy fern ecophysiology (Fig. 2).
Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca.
15 km NW of its larger sister island, Tahiti.
Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
At the time of the study, a total of 18 filmy fern species were known from Moorea (Murdock and Smith, 2003; Ranker et al., 2005; Nitta et al., 2011b; Florence, in press), including multiple terrestrial, saxicolous, and epiphytic species.
The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages (Pryer et al., 2001), are represented by similar numbers of species.
We follow the taxonomic system of Ebihara et al. (2006).

***Field survey***---We surveyed filmy ferns on Moorea as part of a larger study including all ferns of Moorea and Tahiti (Nitta et al., 2016).
During this study, 17 sampling sites were established from ca.
200 m to 1200 m primarily along trails to the summits of three large peaks on Moorea: Mt. Rotui (899 m), Mt. Mouaputa (880 m), and Mt. Tohiea (1206 m; Fig. 2).
At each site, sporophytes were sampled in a single 10 m *×* 10 m plot, and gametophytes sampled in two to three 50 cm *×* 50 cm subplots located within the sporophyte plot.
Epiphytes were sampled to a maximum height of ca.
2 m on their host plants.
Unlike many other fern species, filmy fern gametophytes have unique morphologies that allow them to be identified to family (e.g., filamentous, ribbon with round gemmae; Nayar and Kaur 1971).
This allowed us to specifically use field-collected filmy fern gametophytes for physiological tests, which we later identified to species using DNA barcoding (see below).
We characterized the elevational range of each species based on the results of the field survey; any species found with gametophytes occurring *\>*200 m above the maximum or below the minimum observed elevational range of sporophytes was considered 'widespread' (i.e., facultatively independent sensu Rumsey and Sheffield $$1996$$).
We scored growth habit of each species based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, i.e., below ca. 500 m), or high elevation epiphyte (mostly occurring in cloud forest, i.e., above ca. 500 m).

***Microclimate***---We characterized the microenvironment experienced by filmy ferns by logging relative humidity (RH) and temperature every 15 min from July 7, 2013 to July 5, 2014 at 16 of the sampling sites (Fig. 2).
To do this we used Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA) with external sensors outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation.
To compare between terrestrial and epiphytic habitats, dataloggers were placed in pairs at each site (one 'epiphytic' datalogger at ca. 1.5 m on a tree, one 'terrestrial' datalogger at ca. 10 cm next to the tree).
During the study period, some dataloggers malfunctioned, presumably due to prolonged exposure to high humidity.
We excluded any days during which one or more dataloggers failed from the final dataset, and some sites are missing data completely for one or both dataloggers.
One terrestrial datalogger (Mt. Tohiea 393 m site) was lacking data for more than half of the survey period, but had data available for the same period from a preliminary survey conducted during the previous year (2012--2013).
We used the previous year's data for this site instead.
The final dataset included 201 days of data from 15 sites (26 dataloggers).
We tested for differences in climatic variables between epiphytic and terrestrial habitats using analysis of covariance (ANCOVA) with each climatic variable in turn as the response variable and elevation as the explanatory variable with habitat (terrestrial vs. epiphytic) as a cofactor.

***Desiccation tolerance***---Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
For sporophytes, 8--16 individuals (one individual = single whole frond including a ca. 2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
All experiments were initiated within 48 h of collection.
Pre-treatment chlorophyll fluorescence (F~v~/F~m~) was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100% RH), and water withheld for either a short (2 d) or long (15 d) interval (Testo and Watkins, 2013).
Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22--25 *◦*C.
Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (--282 MPa), Mg(NO~3~)~2~ (--86 MPa), and NaCl (--38 MPa).
Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible; we therefore used the same treatment for all gametophytes (2 d at --86 MPa), and verified species of each individual after the experiment using the DNA barcode approach described below.
Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
Following the desiccation treatment, plants were rewetted using water from a local stream.
Plants were allowed to rehydrate, and chlorophyll fluorescence was again measured at 0*.*5 h, 24 h, and 48 h following rewetting.
Ability to recover from desiccation was quantified by comparing pre- and post-treatment chlorophyll fluorescence yield (Watkins et al., 2007).

***Light responses***---Rapid light response curves were constructed for each species by measuring photosynthetic yield at gradually increasing levels of photosynthetically active actinic light (400--700 nm) with the Light Curve function of the mini-PAM portable chlorophyll fluorometer (Walz Gmbh, Effeltrich, Germany) as described by the manufacturer.
Briefly, the maximum yield (F~m~) was first measured in the absence of actinic light.
Next, plants were equilibrated to the actinic light for 30 s, and photosynthetic yield (∆F/F~m~) measured with a 0*.*8 s saturating pulse.
This was repeated for each photosynthetic photon flux density (PPFD) level, up to ca.
500 µmol·m^--2^ ·s^--1^.
Response curves were fitted using the equation *y* = *A*(1 --*e^-kx^*), where *y* is relative electron transport rate of photosystem II (RETR), *x* is PPFD, *A* is the asymptote of the curve, and *k* is a slope parameter; outliers at high PPFD likely to distort the curve were discarded (Proctor, 2012).
The PPFD at 95% saturation of RETR (PPFD~95%~) and maximum RETR (ETR~max~) were then calculated from the curve.
Light responses were measured in 8--16 field-collected sporophyte individuals for each species, and on single gametophyte individuals in the field, which were later identified to species using the DNA barcodes.

***DNA barcoding***---We used DNA barcoding to identify filmy fern gametophytes to species as described in Nitta et al. (2016).
Briefly, DNA extraction was performed using a modified CTAB protocol (Doyle and Doyle, 1987) or the Plant Mini DNEasy kit (Qiagen, Valencia, California, USA) following the manufacturer's protocol.
Chloroplast *rbcL* and *trnH--psbA* were amplified and sequenced using primers and protocols following Nitta et al. (2016).
For each marker, a reference library was constructed from sporophyte sequences using the makeblastDB command in BLAST (Altschul et al., 1997).
Field-collected gametophytes were identified by a local BLAST query against the reference library.
Gametophytes matching *\>*99.5% with a single reference sequence and no others were identified as that species; those that did not match any reference sequence or matched *\>*99.5% with multiple reference sequences were not identified and excluded from further analysis.

***Statistical analysis and phylogenetic comparative methods***---All statistical analyses were done in R v.3.2.1 (R Core Team, 2015).
Datasets and R scripts used in this study are available on Dryad (\#\#\#\#\#\#).

@@

>>>>

---
title: ""
bibliography: references.bib
mainfont: Roboto-Regular.ttf
fontsize: 12pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    pandoc_args: [ "--csl", "journal-of-plant-research.csl"]
    includes:
      in_header: journal-of-plant-research.sty
header-includes: 
  \usepackage{float}
  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
editor_options: 
  markdown: 
    wrap: sentence
---

<!-- to move to the next line without starting a new paragraph, 
add two spaces after last word, e.g., Atsushi Ebihara__ -->
**Corresponding author**  
Joel H. Nitta  
Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo, 113-0032, Japan  
Phone: +81-  
Fax: +81-  
joelnitta@gmail.com

**Membership of the Botanical Society of Japan**  
YES (Joel H. Nitta)

**Subject area**  
(1)	Taxonomy/Phylogenetics/Evolutionary Biology

**Number of tables, black-and-white figures, and color figures.**  
No tables / 2 black-and-white figure / 1 color figure

<!-- See https://stackoverflow.com/questions/24672111/how-to-add-a-page-break-in-word-document-generated-by-rstudio-markdown/40474753
for inserting page breaks when rendering to MS word -->
#####
\newpage

# Intergenerational Niche Differentiation in Filmy Ferns (Hymenophyllaceae)

Joel H. Nitta^1,4^, James E. Watkins, Jr.^2^, N. Michele Holbrook^1^, Ravahere Taputuarai^3^, Tristan W. Wang^1^, and Charles C. Davis^1^

^1^Department of Organismic and Evolutionary Biology and Harvard University Herbaria, Harvard University, Cambridge, Massachusetts 02138 USA

^2^Department of Biology, Colgate University, 13 Oak Drive, Hamilton, New York 13346 USA

^3^Association Te Rau Ati Ati a Tau a Hiti Noa Tu, Mahina, Tahiti, French Polynesia

^4^Current address: Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo, 113-0032, Japan

Author for correspondence: Joel Nitta (joelnitta\@gmail.com)

ORCID ID numbers  
- Joel H. Nitta: \textnormal{https://orcid.org/0000-0003-4719-7472}  
- N. Michele Holbrook: \textnormal{https://orcid.org/0000-0003-3325-5395}  
- Charles C. Davis: \textnormal{https://orcid.org/0000-0001-8747-1101}

### Abstract
Desiccation tolerance was a key trait allowing plants to colonize land.
However, little is known about the transition from desiccation tolerant non-vascular plants to desiccation sensitive vascular ones.
Filmy ferns (Hymenophyllaceae) present a useful system to study how water-stress strategies differ between non-vascular and vascular stages within a single organism, as they have vascularized sporophytes and nonvascular gametophytes that are each capable of varying degrees of desiccation tolerance.
We surveyed sporophytes and gametophytes of 22 species of filmy ferns on Moorea (French Polynesia) and used chlorophyll fluorescence to measure desiccation tolerance and light responses.
We sequenced plastid *rbcL* and *trnHpsbA* to identify field-collected gametophytes, and conducted phylogenetically informed analyses to identify differences in physiology between life stages and growth habits. 
Gametophytes had similar or less desiccation tolerance and lower photosynthetic optima than sporophytes.
Epiphytes were more tolerant of desiccation than terrestrial species.
Despite their lack of greater physiological tolerances, gametophytes of several species occurred over a wider elevational range than conspecific sporophytes.
Our results show that filmy fern gametophytes and sporophytes differ in their physiology and niche requirements, and point to the importance of microhabitat in shaping the evolution of water-use strategies in vascular plants.

**Keywords**: chlorophyll fluorescence; desiccation tolerance; DNA barcoding; French Polynesia; gametophyte; Hymenophyllaceae

#####
\newpage

## Introduction

@@

Ferns represent an important, understudied group of tracheophytes for investigating the transition from desiccation tolerance to desiccation avoidance in land plants.
Ferns are the only major lineage among land plants with sporophytes and gametophytes that are capable of growing independently from each other.
These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (i.e., they regulate their internal water content), whereas fern gametophytes are much smaller (often <1 cm), lack vascular tissue, and are poikilohydric.
Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT [@WatkinsJr.2007c; @Pittermann2013].
Thus, ferns present a unique opportunity to observe the transition from desiccation tolerance to desiccation avoidance across independently growing life phases within a single organism.

Filmy ferns (Hymenophyllaceae) represent a particularly interesting case for studying the ecological and evolutionary significance of vegetative DT across life stages in ferns.
Filmy ferns are a large family (ca. 430 spp.; PPG I, 2016) of primarily tropical, leptosporangiate ferns named for their remarkably thin leaf laminae, which are usually only a single cell layer in thickness and lack cuticle or stomata (Fig. 1A--D).
Filmy fern sporophytes are an exception to the 'typical' fern sporophyte described above because they have reverted to poikilohydry: their thin laminae allow for passive, rapid uptake and loss of moisture, thereby maintaining nearly constant equilibrium with atmospheric humidity [@Raven1999].
Many species of epiphytic filmy ferns have extremely reduced root systems [@Schneider2000], relying completely upon absorption through leaf laminae for water [@Shreve1911], although terrestrial species possess true roots and use ground water [@Dubuisson2011].
As would be expected from their poikilohydric nature, DT of varying degrees has been reported in sporophytes of many filmy fern species (see references below).

Filmy fern sporophytes encompass a broad diversity of growth forms (terrestrial, saxicolous, hemi-epiphytic, and holoepiphytic) and morphologies, with rhizome and leaf sizes varying over an order of magnitude [@Hennequin2008; @Nitta2009; @Dubuisson2013].
Filmy fern gametophytes are similarly diverse, with morphotypes including filamentous, ribbon, and a combination of these two types (Fig. 1E--H).
While there are no quantitative data about DT in filmy fern gametophytes available, DT has been reported in gametophytes from several other fern lineages [@WatkinsJr.2007a; @WatkinsJr.2012; @Pittermann2013], and may be expected to occur in filmy ferns as well.
The possibility for varying degrees of vegetative DT in both stages of the life cycle of filmy ferns makes them a potentially useful system to compare how DT is utilized between haploid, non-vascular and diploid, vascular life stages.
Furthermore, the diversity of growth forms and habitats in filmy ferns makes them an ideal group for a comparative study of water-stress strategies (i.e., desiccation tolerance vs. desiccation avoidance) and the correlation of these strategies with ecological niche.

The degree of DT in filmy fern sporophytes occupying different niches has been relatively well studied.
In a series of elegant experiments on Jamaican filmy ferns, [@Shreve1911] demonstrated that water-use strategies differ between species occupying contrasting habitats (i.e., terrestrial, low elevation epiphyte, and high elevation epiphyte).
Proctor [-@Proctor2003; -@Proctor2012] investigated water relations and the ability to recover from desiccation in sporophytes of several filmy ferns, and found a general correlation between high DT and adaptation for high light levels in widespread species vs. low DT and adaptation to low light levels in species occupying more sheltered habitats.
In the only study to our knowledge including in situ physiological observations of hymenophyllaceous gametophytes, @Johnson2000 found that gametophytes of *Vandenboschia speciosa* (Willd.) G.
Kunkel (= *Trichomanes speciosum* Willd.) are adapted for extremely low light environments, and operate at lower light levels than conspecific sporophytes.
@Parra2009 and @Saldana2013 investigated the relationship between physiology (photosynthetic rates and water relations) and vertical stratification in several epiphytic filmy fern species (primarily genus *Hymenophyllum*) along tree trunks in Chile.
They found that species occurring at greater heights in the trees tended to have greater DT, higher photosynthetic capacity (A~max~), and lower rates of evapotranspiration.
A recent study using two species of *Hymenophyllum* demonstrated that DT in filmy ferns likely operates via a constitutive mechanism, which is more similar to that of bryophytes than other drought-tolerant vascular plants, thus allowing filmy ferns to adjust to dry conditions extremely rapidly [@GarcesCea2014].

Despite this considerable interest in filmy fern ecophysiology, with the sole exception of @Johnson2000, no studies have compared physiology between gametophyte and sporophyte phases, and we are aware of none that have done so across multiple species.
Furthermore, to our knowledge, no physiological studies of filmy ferns have accounted for their phylogenetic relationships, and we lack an understanding of how physiological strategies have evolved in this clade.
In addition to DT, photosynthetic optima (adaptation to different light levels) are likely to be important in determining the niches of life stages and species of filmy ferns [@Proctor2003; @Parra2015; @Flores-Bavestrello2016].
Terrestrial and epiphytic species experience different levels of photosynthetic radiation and desiccation, and light intensity varies with height along the host plant in epiphytes  [@Benzing1990; @WatkinsJr.2009].
In a recent study (Testo et al., unpublished), the combined effect of desiccation and light stress was found to have a greater impact on filmy fern physiological performance than either of these stressors alone, suggesting the need to integrate studies of DT with photosynthesis to gain a full picture of physiological adaptation in filmy ferns.
Here, we leverage a recently developed DNA barcoding system [@Nitta2017] to test several hypotheses related to DT and photosynthesis in a comparative phylogenetic framework including both filmy fern sporophytes and gametophytes: 1.) Epiphytic filmy ferns have higher DT and photosynthetic optima than non-epiphytic species.
2.) Amongst epiphytic filmy ferns, those occurring at high elevation have higher DT and photosynthetic optima than those at low elevation.
3.) For filmy ferns with gametophytes that occur over a wider range than sporophytes, the gametophyte has higher DT and photosynthetic optima than the sporophyte.

<!-- 
## Materials and methods

### Study system

We selected the island of Moorea, French Polynesia (17*◦*32' S, 149*◦*50' W) for our survey of filmy fern ecophysiology (Fig. 2).
Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca.
15 km NW of its larger sister island, Tahiti.
Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
At the time of the study, a total of 18 filmy fern species were known from Moorea (Murdock and Smith, 2003; Ranker et al., 2005; Nitta et al., 2011b; Florence, in press), including multiple terrestrial, saxicolous, and epiphytic species.
The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages (Pryer et al., 2001), are represented by similar numbers of species.
We follow the taxonomic system of Ebihara et al. (2006).

### Field survey

We surveyed filmy ferns on Moorea as part of a larger study including all ferns of Moorea and Tahiti (Nitta et al., 2016).
During this study, 17 sampling sites were established from ca.
200 m to 1200 m primarily along trails to the summits of three large peaks on Moorea: Mt. Rotui (899 m), Mt. Mouaputa (880 m), and Mt. Tohiea (1206 m; Fig. 2).
At each site, sporophytes were sampled in a single 10 m *×* 10 m plot, and gametophytes sampled in two to three 50 cm *×* 50 cm subplots located within the sporophyte plot.
Epiphytes were sampled to a maximum height of ca.
2 m on their host plants.
Unlike many other fern species, filmy fern gametophytes have unique morphologies that allow them to be identified to family (e.g., filamentous, ribbon with round gemmae; Nayar and Kaur 1971).
This allowed us to specifically use field-collected filmy fern gametophytes for physiological tests, which we later identified to species using DNA barcoding (see below).
We characterized the elevational range of each species based on the results of the field survey; any species found with gametophytes occurring *\>*200 m above the maximum or below the minimum observed elevational range of sporophytes was considered 'widespread' (i.e., facultatively independent sensu Rumsey and Sheffield $$1996$$).
We scored growth habit of each species based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, i.e., below ca. 500 m), or high elevation epiphyte (mostly occurring in cloud forest, i.e., above ca. 500 m).

### Microclimate

We characterized the microenvironment experienced by filmy ferns by logging relative humidity (RH) and temperature every 15 min from July 7, 2013 to July 5, 2014 at 16 of the sampling sites (Fig. 2).
To do this we used Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA) with external sensors outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation.
To compare between terrestrial and epiphytic habitats, dataloggers were placed in pairs at each site (one 'epiphytic' datalogger at ca. 1.5 m on a tree, one 'terrestrial' datalogger at ca. 10 cm next to the tree).
During the study period, some dataloggers malfunctioned, presumably due to prolonged exposure to high humidity.
We excluded any days during which one or more dataloggers failed from the final dataset, and some sites are missing data completely for one or both dataloggers.
One terrestrial datalogger (Mt. Tohiea 393 m site) was lacking data for more than half of the survey period, but had data available for the same period from a preliminary survey conducted during the previous year (2012--2013).
We used the previous year's data for this site instead.
The final dataset included 201 days of data from 15 sites (26 dataloggers).
We tested for differences in climatic variables between epiphytic and terrestrial habitats using analysis of covariance (ANCOVA) with each climatic variable in turn as the response variable and elevation as the explanatory variable with habitat (terrestrial vs. epiphytic) as a cofactor.

### Desiccation tolerance

Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
For sporophytes, 8--16 individuals (one individual = single whole frond including a ca. 2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
All experiments were initiated within 48 h of collection.
Pre-treatment chlorophyll fluorescence (F~v~/F~m~) was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100% RH), and water withheld for either a short (2 d) or long (15 d) interval (Testo and Watkins, 2013).
Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22--25 *◦*C.
Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (--282 MPa), Mg(NO~3~)~2~ (--86 MPa), and NaCl (--38 MPa).
Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible; we therefore used the same treatment for all gametophytes (2 d at --86 MPa), and verified species of each individual after the experiment using the DNA barcode approach described below.
Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
Following the desiccation treatment, plants were rewetted using water from a local stream.
Plants were allowed to rehydrate, and chlorophyll fluorescence was again measured at 0*.*5 h, 24 h, and 48 h following rewetting.
Ability to recover from desiccation was quantified by comparing pre- and post-treatment chlorophyll fluorescence yield (Watkins et al., 2007).

### Light responses

Rapid light response curves were constructed for each species by measuring photosynthetic yield at gradually increasing levels of photosynthetically active actinic light (400--700 nm) with the Light Curve function of the mini-PAM portable chlorophyll fluorometer (Walz Gmbh, Effeltrich, Germany) as described by the manufacturer.
Briefly, the maximum yield (F~m~) was first measured in the absence of actinic light.
Next, plants were equilibrated to the actinic light for 30 s, and photosynthetic yield (∆F/F~m~) measured with a 0*.*8 s saturating pulse.
This was repeated for each photosynthetic photon flux density (PPFD) level, up to ca.
500 µmol·m^--2^ ·s^--1^.
Response curves were fitted using the equation *y* = *A*(1 --*e^-kx^*), where *y* is relative electron transport rate of photosystem II (RETR), *x* is PPFD, *A* is the asymptote of the curve, and *k* is a slope parameter; outliers at high PPFD likely to distort the curve were discarded (Proctor, 2012).
The PPFD at 95% saturation of RETR (PPFD~95%~) and maximum RETR (ETR~max~) were then calculated from the curve.
Light responses were measured in 8--16 field-collected sporophyte individuals for each species, and on single gametophyte individuals in the field, which were later identified to species using the DNA barcodes.

### DNA barcoding

We used DNA barcoding to identify filmy fern gametophytes to species as described in Nitta et al. (2016).
Briefly, DNA extraction was performed using a modified CTAB protocol (Doyle and Doyle, 1987) or the Plant Mini DNEasy kit (Qiagen, Valencia, California, USA) following the manufacturer's protocol.
Chloroplast *rbcL* and *trnH--psbA* were amplified and sequenced using primers and protocols following Nitta et al. (2016).
For each marker, a reference library was constructed from sporophyte sequences using the makeblastDB command in BLAST (Altschul et al., 1997).
Field-collected gametophytes were identified by a local BLAST query against the reference library.
Gametophytes matching *\>*99.5% with a single reference sequence and no others were identified as that species; those that did not match any reference sequence or matched *\>*99.5% with multiple reference sequences were not identified and excluded from further analysis.

### Statistical analysis and phylogenetic comparative methods

All statistical analyses were done in R v.3.2.1 (R Core Team, 2015).
Datasets and R scripts used in this study are available on Dryad (\#\#\#\#\#\#).

@@

-->

## References
<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->

<div id="refs"></div>
